---
title: "Billboard Genres"
output: html_notebook
---
```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(clock)
library(lubridate)
```

```{r}
billboard <- read_csv("../data/billboard_2000_2018_spotify_lyrics.csv")
```
```{r}
#Removing the video link, simple title column.
billboard <- billboard |> 
  select(-c(video_link, simple_title, spotify_link, spotify_id, lyrics, analysis_url, genre, mode, key, time_signature))
```

```{r}
#Changing columns to numeric
billboard <- billboard |> 
  mutate(across(c(energy, instrumentalness, liveness, tempo, speechiness, acousticness, danceability, duration_ms, loudness, valence), as.numeric))
```

```{r}
#Removing the data with any missing information.
billboard <- 
  billboard |> 
  drop_na()
```

```{r}
#Creating a column with just the months & numbered weeks
billboard <- billboard |> 
  mutate(date = as.Date(date, format = "%m/%d/%y"),
         month = month(date, label = TRUE),
         week_number = week(date))
```

```{r}
#creating a column that will have the season a song joined the charts.
billboard <- billboard %>%
  mutate(seasons = case_when(
    month %in% c("Dec", "Jan", "Feb") ~ "Winter",
    month %in% c("Mar", "Apr", "May") ~ "Spring",
    month %in% c("Jun", "Jul", "Aug") ~ "Summer",
    month %in% c("Sep", "Oct", "Nov") ~ "Fall"
  ))
```

```{r}
billboard <- 
  billboard |> 
  mutate(number_one = peak_pos == 1)
```

title, energy, liveness, speechiness, ascousticness, instrumentalness, danceability, valence

select(-c(year, peak_pos, weeks, rank, duration_ms, last_pos, mode, time_signature, key, tempo, loudness, date, main_artist, change, genre, broad_genre, month, week_number, seasons, artist)) 

```{r}
top_songs <- billboard  |> 
  group_by(year, broad_genre)  |> 
  slice_max(order_by = weeks, n = 1)  |> 
  ungroup() 

billboard_pivoted <-
  billboard |> 
  filter(title %in% (top_songs |> pull(title))) |> 
  select(title,year,main_artist, broad_genre, energy, liveness, speechiness, acousticness, instrumentalness, danceability, valence) |> 
pivot_longer(cols =-c("title", "year", "broad_genre","main_artist"), names_to = "variable", values_to = "value")  
  
```

ARTISTS
```{r}
billboard |> 
  filter(weeks >= 14 & rank <= 50) |> 
  count(main_artist) |> 
  arrange(desc(n)) |> 
  filter(n > 1) |> 
  mutate(median_of_n = median(n)) 
  
  
```

```{r}
#artist with more than 3 songs that charted at least 14 weeks and ranked lower than 50. 
top_artists <-
billboard |> 
  filter(weeks >= 14 & rank <= 50) |> 
   group_by(main_artist)  |> 
  filter(n() > 3)  |> 
  arrange(desc(n()))

top_artists
```

```{r}
billboard |> 
  filter(change == "Hot Shot Debut")
```
```{r}
billboard |> 
  filter(change == "New")
```
```{r}
billboard |> 
  filter(change == "Re-Entry")
```






```{r}
# billboard <- billboard |> 
#   filter(weeks >= 14, rank <= 50)
```

```{r}
# write.csv(billboard, "../data/billboard_genre\\.csv", row.names = FALSE)
```
```{r}
# write.csv(billboard_pivoted, "../data/billboard_pivot\\.csv", row.names = FALSE)
```


```{r}
# billboard |> 
#   select(where(is.numeric), -year, -peak_pos, -weeks, -rank, -duration_ms, -last_pos, -mode, -time_signature, -key, -tempo, -loudness) |> 
#   colMeans() |> 
#   enframe(name = "Variable", value = "Mean")  |> 
#   ggplot(aes(x = Variable, y = Mean)) +
#   geom_col() +
#   labs(title = "Column Means", x = "Variables", y = "Mean Value")

```
REGRESSIONS

```{r}
logistic_model <- glm(number_one ~ weeks + rank, data = billboard, family = "binomial")

summary(logistic_model)
```
energy, liveness, speechiness, acousticness, instrumentalness, danceability, valence
```{r}
#Logistic models that checks how energy, liveness, speechiness, acousticness, instrumentalness, danceability, valence may determine if a song reaches the number one spot. 
logistic_model <- glm(number_one ~ energy + liveness + speechiness + acousticness + instrumentalness + danceability + valence, data = billboard, family = "binomial")

summary(logistic_model)
```

```{r}

```

```{r}

```







```{r}
# billboard_weeks_lm <- lm("weeks ~ energy", data = billboard)
# summary(billboard_weeks_lm)
```

```{r}
# reduced_model <- lm("weeks ~ energy", data = billboard)
# full_model <- lm("weeks ~ energy + liveness", data = billboard)
# 
# anova(reduced_model, full_model)
```












































